---
title: "Data Analysis Script"
format: html
---

# Exploration

### Packages Install

```{r packages}
install.packages("tidyverse")
```

### Import csv file

```{r}
read_csv(file = "../data/cleaned_data/cleaned_all_star_data.csv") -> nba_all_stars
```

### Exploration Questions

1.  How has the number of three pointers attempted changed over time? The vertical line represents the year Stephen Curry entered the NBA

```{r}
nba_all_stars |>
  group_by(year) |>
  summarize(avg_fg3a = mean(fg3a, na.rm = TRUE)) -> avg_3_atempt

avg_3_atempt |>
  ggplot(mapping = aes(x = year,
                       y = avg_fg3a)) +
  geom_point(color = "blue") +
  geom_smooth(color = "black", alpha = 0.25) +
  geom_vline(xintercept = 2009, color = "gold")
# saving result to output file
ggsave(filename = "../output/question_1.png")
```

2.  Top ten players who scored the highest average points in years they were an all star

```{r}
nba_all_stars |>
  group_by(first, last) |>
  summarize(avg_pts = mean(pts, na.rm = TRUE)) |>
  arrange(desc(avg_pts)) |>
  head(n = 10) -> question_2
# saving results to output file
write_csv(x = question_2,
          file = "../output/question_2.csv")
  
```

3.  How does average points per game correlate with the average free throws attempted?

```{r}
nba_all_stars |>
  ggplot(mapping = aes(x = pts,
                       y = fta)) +
  geom_point() + 
  geom_smooth(method = "lm")
# saving results to output file
ggsave(filename = "../output/question_3.png")
```

4.  Which team has contributed the most all stars and the least all stars?

```{r}
nba_all_stars |>
  group_by(team) |>
  summarise(count = n()) |>
  arrange(desc(count)) -> question_4_a
# saving results of part a to output file
write_csv(x = question_4_a,
          file = "../output/question_4_a.csv")

nba_all_stars |>
  group_by(team) |>
  summarise(count = n()) |>
  arrange(count) -> question_4_b
# saving resluts of part b to output file
write_csv(x = question_4_b,
          file = "../output/question_4_b.csv")
```

5.  What is the correlation between assists and turnovers?

```{r}
nba_all_stars |>
  ggplot(mapping = aes(x = ast,
                       y = turnover)) +
  geom_point(color = "paleturquoise4") +
  geom_smooth(method = "lm", color = "hotpink")
#saving results to output file
ggsave(filename = "../output/question_5.png")
```

6.  Which players have the most all star selections?

```{r}
nba_all_stars |>
  group_by(first, last) |>
  summarize(count = n()) |>
  arrange(desc(count)) -> question_6
#saving result to output file
write_csv(x = question_6,
          file = "../output/question_6.csv")
  

```

### Final data frame

```{r}
# creates the final data frame to be posted 
write_csv(x = nba_all_stars,
          file = "../output/final_data.csv")
```
